Equalizer 2\\Convergence of a Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
grammar
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@52 db7c848b-7755-4df2-b763-8796797afa52

Sort-first (\fig{fig:loadeq}) and sort-last load balancing is the most obvious
optimization for these parallel rendering modes. Our load equalizer is fully
transparent for application developers, that is, it uses a reactive approach
based on past rendering times. This assumes a reasonable frame-to-frame
coherency. Our implementation stores a 2D or 1D grid of the load, mapping the
load of each channel. The load is stored in normalized 2D/1D coordinates using
$\frac{time}{area}$ as the load, the contributing source channels are organized
in a binary tree, and then the algorithm balances the two branches of each level
by equalizing the integral over the area on each side.

@@

The DFR equalizer (\fig{fig:dfr}) provides a functionality similar to dynamic
video resizing \cite{MBDM:97}, that is, it maintains a constant framerate by
adapting the rendering resolution of a fill-limited application. In \textsf{Equalizer},
this works by rendering into a source channel (typically on a FBO) separate to
the destination channel, and then scaling the rendering during the transfer
(typically through an on-GPU texture) to the destination channel. The DFR
equalizer monitors the rendering performance and accordingly adapts the
resolution of the source channel and zoom factor for the source to destination
transfer. If the performance and source channel resolutions allows, this will
not only subsample, but also supersample the destination channel to reduce
aliasing artefacts.

@@

In full synchronization, all threads always execute the same frame, that is, the
render threads are unlocked after \textsf{Node::frameStart}, and the node is
blocked for all render threads to finish the frame before executing
\textsf{Node::frameFinish}. This allows the render threads to read shared data
from all their operations, but provides the slowest performance.

In draw synchronization, the node thread and all render threads are synchronized
for all \textsf{frameDraw} operations, that is, \textsf{Node::frameFinish} is
executed after the last channel is done drawing. This allows the render threads
to read shared data during their draw operation, but not during
compositing. Since compositing is typically independent of the rendered data,
this is the default mode. This mode allows to overlap compositing with data
synchronization on multi-GPU machines.

@@

To our knowledge, all parallel rendering systems have the focal plane coincide
with the physical display surface. For better viewing comfort, we introduce a
new dynamic focus mode, where the application defines the distance of the focal
plane from the observer, based on the current view direction of the
user. Initial experiments show that this is particularly effective for objects
placed within the immersive space, that is, in front of all display segments.

@@

A client opens a listening multicast RSP connection. Similar to TCP, this
listener will create one connected RSP connection for each remote member in the
multicast group, that is, on each node $n$ RSP connection instances exist, $1$
for the local listener and $n-1$ for all other nodes. A multicast group
therefore looks like a fully connected set of TCP connections between $n$ nodes
for the user.

@@

Congestion control is necessary to optimize bandwidth usage. While TCP uses the
well-known AIMD (additive increase, multiplicative decrease) algorithm, we have
chosen a more aggressive congestion control algorithm of additive increase and
decrease. This has proven experimentally to be more optimal: UDP is often
rate-limited by switches, that is, packets are discarded regularly and not
exceptionally. Only slowly backing of the current send rate helps to stay close
to this limit. Furthermore, our RSP traffic is limited to the local subnet,
making cooparation between multiple data stream less of an issue. Send rate
limiting uses a bucket algorithm, where over time the bucket fills with send
credits, from which sends are substracted. If there are no available credits,
the sender sleeps until sufficient credits are available.

@@

\begin{compactdesc}
\item[Static] The object is not versioned nor buffered. The instance data is
  serialized whenever a new slave instance is mapped. No additional data is
  stored.
\item[Instance] The object is versioned and buffered. The instance and delta
  data are identical, that is, only instance data is serialized. Previous
  instance data is saved to be able to map old versions.
\item[Delta] The object is versioned and buffered. The delta data is typically
  smaller than the instance data. The delta data is transmitted to slave
  instances for synchronization. Previous instance and delta data is saved to be
  able to map and sync old versions.
\item[Unbuffered] The object is versioned and unbuffered. No data is stored, and
  no previous versions can be mapped. No old versions can be mapped and no
  additional data is stored.
\end{compactdesc}

@@

\begin{compactitem}
\item The master instance of the object generates new versions for all
  slaves. These versions are continuous. It is possible to commit on slave
  instances, but special care has to be taken to handle possible
  conflicts.
\item Slave instance versions can only be advanced, that is, \textsf{sync(
  version )} with a version smaller than the current version will fail.
\item Newly mapped slave instances are mapped to the oldest available
  version by default, or to the version specified when calling
  \textsf{mapObject}.
\end{compactitem}

@@

Unsurprisingly, TCP over the faster InfiniBand link outperforms the cheaper ten
gigabit ethernet link by more than a factor of two. Unexpectedly, the native
RDMA connection performs worse, even though it outperforms IPoIB in a simple
peer-to-peer connection benchmark. This needs further investigations, but we
suspect the abstraction of a byte stream connection chosen by \textsf{Collage}
for historical reasons is not will suited for remote DMA semantics, that is,
one needs to design the network API around zero-copy semantics with managed
memory for modern high-speed transports. Both Infiniband connections show
significant measurement jitter.

>>>>

Sort-first (\fig{fig:loadeq}) and sort-last load balancing is the most obvious
optimization for these parallel rendering modes. Our load equalizer is fully
transparent for application developers; that is, it uses a reactive approach
based on past rendering times. This assumes a reasonable frame-to-frame
coherency. Our implementation stores a 2D or 1D grid of the load, mapping the
load of each channel. The load is stored in normalized 2D/1D coordinates using
$\frac{time}{area}$ as the load, the contributing source channels are organized
in a binary tree, and then the algorithm balances the two branches of each level
by equalizing the integral over the area on each side.

@@

The DFR equalizer (\fig{fig:dfr}) provides a functionality similar to dynamic
video resizing \cite{MBDM:97}; that is, it maintains a constant framerate by
adapting the rendering resolution of a fill-limited application. In
\textsf{Equalizer},
this works by rendering into a source channel (typically on a FBO) separate to
the destination channel, and then scaling the rendering during the transfer
(typically through an on-GPU texture) to the destination channel. The DFR
equalizer monitors the rendering performance and accordingly adapts the
resolution of the source channel and zoom factor for the source to destination
transfer. If the performance and source channel resolutions allows, this will
not only subsample, but also supersample the destination channel to reduce
aliasing artefacts.

@@

In full synchronization, all threads always execute the same frame; that is, the
render threads are unlocked after \textsf{Node::frameStart}, and the node is
blocked for all render threads to finish the frame before executing
\textsf{Node::frameFinish}. This allows the render threads to read shared data
from all their operations, but provides the slowest performance.

In draw synchronization, the node thread and all render threads are synchronized
for all \textsf{frameDraw} operations; that is, \textsf{Node::frameFinish} is
executed after the last channel is done drawing. This allows the render threads
to read shared data during their draw operation, but not during
compositing. Since compositing is typically independent of the rendered data,
this is the default mode. This mode allows to overlap compositing with data
synchronization on multi-GPU machines.

@@

To our knowledge, all parallel rendering systems have the focal plane coincide
with the physical display surface. For better viewing comfort, we introduce a
new dynamic focus mode, where the application defines the distance of the focal
plane from the observer, based on the current view direction of the
user. Initial experiments show that this is particularly effective for objects
placed within the immersive space; that is, in front of all display segments.

@@

A client opens a listening multicast RSP connection. Similar to TCP, this
listener will create one connected RSP connection for each remote member in the
multicast group; that is, on each node $n$ RSP connection instances exist, $1$
for the local listener and $n-1$ for all other nodes. A multicast group
therefore looks like a fully connected set of TCP connections between $n$ nodes
for the user.

@@

Congestion control is necessary to optimize bandwidth usage. While TCP uses the
well-known AIMD (additive increase, multiplicative decrease) algorithm, we have
chosen a more aggressive congestion control algorithm of additive increase and
decrease. This has proven experimentally to be more optimal: UDP is often
rate-limited by switches; that is, packets are discarded regularly and not
exceptionally. Only slowly backing of the current send rate helps to stay close
to this limit. Furthermore, our RSP traffic is limited to the local subnet,
making cooparation between multiple data stream less of an issue. Send rate
limiting uses a bucket algorithm, where over time the bucket fills with send
credits, from which sends are substracted. If there are no available credits,
the sender sleeps until sufficient credits are available.

@@

\begin{compactdesc}
\item[Static] The object is not versioned nor buffered. The instance data is
  serialized whenever a new slave instance is mapped. No additional data is
  stored.
\item[Instance] The object is versioned and buffered. The instance and delta
  data are identical; that is, only instance data is serialized. Previous
  instance data is saved to be able to map old versions.
\item[Delta] The object is versioned and buffered. The delta data is typically
  smaller than the instance data. The delta data is transmitted to slave
  instances for synchronization. Previous instance and delta data is saved to be
  able to map and sync old versions.
\item[Unbuffered] The object is versioned and unbuffered. No data is stored, and
  no previous versions can be mapped. No old versions can be mapped and no
  additional data is stored.
\end{compactdesc}

@@

\begin{compactitem}
\item The master instance of the object generates new versions for all
  slaves. These versions are continuous. It is possible to commit on slave
  instances, but special care has to be taken to handle possible
  conflicts.
\item Slave instance versions can only be advanced; that is, \textsf{sync(
  version )} with a version smaller than the current version will fail.
\item Newly mapped slave instances are mapped to the oldest available
  version by default, or to the version specified when calling
  \textsf{mapObject}.
\end{compactitem}

@@

Unsurprisingly, TCP over the faster InfiniBand link outperforms the cheaper ten
gigabit ethernet link by more than a factor of two. Unexpectedly, the native
RDMA connection performs worse, even though it outperforms IPoIB in a simple
peer-to-peer connection benchmark. This needs further investigation, but we
suspect the abstraction of a byte stream connection chosen by \textsf{Collage}
for historical reasons is not well suited for remote DMA semantics; that is,
one needs to design the network API around zero-copy semantics with managed
memory for modern high-speed transports. Both Infiniband connections show
significant measurement jitter.
